Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-11 docs/proposals/NSF/ProjDesc/ProjDesc.tex
working on proposal

%---------------------------------------------------------------------------------------------------%
%% PACKAGES
%\usepackage{fullpage}
%\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx}
\usepackage{natbib, hyperref, url}
%\usepackage[justification=RaggedRight]{caption}
\usepackage{outlines,enumitem}
\usepackage{wrapfig, todonotes}
\usepackage{titlesec}
\usepackage{fancyhdr}
%---------------------------------------------------------------------------------------------------%

@@

%---------------------------------------------------------------------------------------------------%
%% TITLE
\begin{document}
\begin{center}
{\large \bf Erosion, Transport, and Dispersion in Granular
and Porous Media} 
\end{center}
%\title{Project Description --
%Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media}

@@

%---------------------------------------------------------------------------------------------------%
\subsection{Transport through eroded media}
\begin{wrapfigure}[]{r}{0.5\textwidth}
  \includegraphics[width=0.5\textwidth]{figs/100b_t100tracer}
  \caption{\label{fig:100tracers} A visualization of the streamlines
  with 200 tracers passing through an eroded geometry with a porosity of
  0.63. The tracers are initialized at the left end of the channel.}
\end{wrapfigure}
Transport through eroded media is a direct consequence of the pore
structure. Qualitative properties of an eroded media, such as channels
of high porosity can be modest in volume fraction, but still be
responsible for a large portion of the flux~\cite{Quaife2018}. To
increase understanding of flow through eroded media, we will simulate
and statistically analyze transport through eroded geometries.

@@

While the tortuosity characterizes the amount of winding of streamlines,
the dispersion characterizes the spreading of the trajectories.
Dispersion can be descried using the first- and second-ensemble moments
of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{}. In Figure~\ref{fig:100dispersion}, we
investigate spreading at several different porosities of a porous
geometry that is initialized with 100 grains.

\begin{wrapfigure}[]{r}{0.5\textwidth}
  \includegraphics[width=0.5\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading,
  characterized with the standard deviation $\sigma_\lambda(t)$ at 6
  different porosities. At early times, the line of slope 1 (black
  curve) indicates a ballistic motion. Then, at later times, the
  dispersion becomes super-dispersive at a rate indicated by the slope
  of the black dashed-dotted line.}
\end{wrapfigure}

@@

We propose to develop a new method that uses the Laplace transform.
Upon applying a Strang splitting, the resulting diffusion equation is
\begin{alignat}{3}
  \pd{c}{t} &= \frac{1}{\Pe} \Delta c, \quad &&\xx \in \Omega, \: t>0, \\
  \pd{c}{\nn} &= 0, &&\xx \in \partial\Omega, \: t>0, \\
  c(\xx,0) &= c_0(\xx), &&\xx \in \Omega.
\end{alignat}
Then, the Laplace transform, $C(\xx,s) = \mathcal{L}[c](\xx,t)$,
satisfies
\begin{align}
  sC(\xx,s) - \frac{1}{\Pe} = c_0(\xx).
\end{align}
This PDE will be solved with high-order accuracy and efficiently by
using fast-multipole-accelerated integral equation methods of the
Beltrami equation~\cite{kro-qua2011}, and the concentration will be
recovered with a inverese Laplace transform~\cite{jos-war2006}.

>>>>

%---------------------------------------------------------------------------------------------------%
%% PACKAGES
%\usepackage{fullpage}
%\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{graphicx, amsmath, amssymb, amsfonts, mathtools, mathrsfs, color}
\usepackage{comment, enumerate, tabularx}
\usepackage{natbib, hyperref, url}
%\usepackage[justification=RaggedRight]{caption}
\usepackage{outlines,enumitem}
\usepackage{wrapfig, todonotes}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{subcaption}
%---------------------------------------------------------------------------------------------------%

@@

%---------------------------------------------------------------------------------------------------%
%% TITLE
\begin{document}
\begin{center}
{\large \bf Erosion, Transport, and Dispersion in Granular
and Porous Media} 
\\
M.J.N.~Moore and Bryan D.~Quaife
\end{center}
%\title{Project Description --
%Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media}

@@

%---------------------------------------------------------------------------------------------------%
\subsection{Transport through eroded media}
\begin{wrapfigure}[]{r}{0.3\textwidth}
  \includegraphics[width=0.3\textwidth]{figs/100b_t100tracer}
  \caption{\label{fig:100tracers} A visualization of the streamlines
  with 200 tracers passing through an eroded geometry with a porosity of
  0.63. The tracers are initialized at the left end of the channel.}
\end{wrapfigure}
Transport through eroded media is a direct consequence of the pore
structure. Qualitative properties of an eroded media, such as channels
of high porosity can be modest in volume fraction, but still be
responsible for a large portion of the flux~\cite{Quaife2018}. To
increase understanding of flow through eroded media, we will simulate
and statistically analyze transport through eroded geometries.

@@

While the tortuosity characterizes the amount of winding of streamlines,
the dispersion characterizes the spreading of the trajectories.
Dispersion can be descried using the first- and second-ensemble moments
of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{}. In Figure~\ref{fig:100dispersion}, we
investigate spreading at several different porosities of a porous
geometry that is initialized with 100 grains.
\begin{wrapfigure}[]{r}{0.5\textwidth}
  \includegraphics[width=0.5\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading,
  characterized with the standard deviation $\sigma_\lambda(t)$ at 6
  different porosities. At early times, the line of slope 1 (black
  curve) indicates a ballistic motion. Then, at later times, the
  dispersion becomes super-dispersive at a rate indicated by the slope
  of the black dashed-dotted line.}
\end{wrapfigure}

The rate of dispersion is known to be directly related to the pore size
distribution~\cite{dea-qua-bir-jua2018}. Therefore, there is a need to
understand how erosion affects the pore size distribution. In
preliminary, we define pore sizes between grains that share an edge of a
Delanuay triangulation with vertices at the grain centers
(Figure~\ref{fig:triangluation}). In Figure~\ref{fig:poreStats}, we plot
the mean and variance of the pore sizes as a function of porosity.

\begin{figure}
\begin{subfigure}[b]{0.33\textwidth}
  \includegraphics*[height=0.8\linewidth]{figs/triangulation_100b100.pdf}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
  \includegraphics*[height = 0.8\linewidth]{figs/gap_mean}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
  \includegraphics*[height=0.8\linewidth]{figs/gap_variance}
\caption{}
\end{subfigure}
  \caption{\label{fig:Eroding100gap_mean_var} (a) The pore sizes are
  defined in terms of a Delanuay triangulation. The effect of erosion on
  (b) the mean and (c) the variance of the pore sizes. The geometry
  initially contains 100 eroding bodies.}
\end{figure}

@@

We propose to develop a new method that uses the Laplace transform.
Upon applying a Strang splitting, the resulting diffusion equation is
\begin{alignat}{3}
  \pd{c}{t} &= \frac{1}{\Pe} \Delta c, \quad &&\xx \in \Omega, \: t>0, \\
  \pd{c}{\nn} &= 0, &&\xx \in \partial\Omega, \: t>0, \\
  c(\xx,0) &= c_0(\xx), &&\xx \in \Omega.
\end{alignat}
Then, the Laplace transform, $C(\xx,s) = \mathcal{L}[c](\xx,t)$,
satisfies
\begin{align}
  sC(\xx,s) - \frac{1}{\Pe}\Delta C(\xx,s) = c_0(\xx), 
    \quad \xx \in \Omega
\end{align}
This PDE will be solved with high-order accuracy and efficiently by
using fast-multipole-accelerated integral equation methods of the Yukawa
equation~\cite{kro-qua2011}, and the concentration will be recovered
with an inverse Laplace transform~\cite{jos-war2006}.
